---
title: "Week 1 In-Class Exercise: Welcome & Modern Toolkit"
subtitle: "GPH-GU 2182 · Statistical Programming in R"
author: "Shawnee King & Yehrim Hwang"
date: "September 2, 2026"
format:
  html:
    toc: true
    number-sections: true
    code-tools:
      source: true
      toggle: false
      caption: "Source Code"
---

::: callout-note
### Instructions

Work in pairs during class.

-   📥 [**Download `week-01-inclass.qmd`**](https://raw.githubusercontent.com/gph-2182/gph-2182.github.io/main/weeks/inclass/week-01-inclass.qmd) to open and edit in RStudio (or use the **`</> Source Code`** button at the top right).
-   Render this Quarto document to HTML and push it to your GitHub repository by the end of class for participation credit.
:::

# Cycle 1: R as a Calculator and Environment Setup

## Basic Arithmetic & Calculations

In the code chunk below, compute the Body Mass Index (BMI) for three study participants:

-   Participant A: Weight $70\text{ kg}$, Height $1.75\text{ m}$
-   Participant B: Weight $85\text{ kg}$, Height $1.80\text{ m}$
-   Participant C: Weight $58\text{ kg}$, Height $1.62\text{ m}$

$$\text{BMI} = \frac{\text{weight in kg}}{(\text{height in meters})^2}$$

```{r bmi-calculations}
# Write your calculations below:
weight_in_kg <- c(70, 85, 58)
height_in_meters <- c(1.75, 1.80, 1.62)
BMI <- weight_in_kg/(height_in_meters)^2
BMI
```

## Exploring Built-in Functions

Compute the mean of the three BMIs you calculated above. Round the result to 2 decimal places using `round()`.

```{r bmi-summary}
# Write your summary code below:
mean(BMI)
round(mean(BMI), 2)
```

## R Packages and Documentation

1.  Load the `r02pro` package (install it in the Console with `install.packages("r02pro")` if you haven't already).
2.  Display the first 6 rows of the `ahp` dataset using the `head()` function.
3.  Open the help file for the `median()` function using `?median`.

```{r r02pro-check}
# Load library and inspect data:
library("r02pro")
head(ahp)
?median
```

------------------------------------------------------------------------

# Cycle 2: Reproducible Reporting with Quarto

## Text Formatting & Markdown

In the space below:

-   Write a short 2-sentence paragraph describing the purpose of reproducible research in public health.
-   Include at least one **bold** phrase and one *italicized* word.
-   Add a bulleted list of two advantages of Quarto over traditional copy-pasting into Word.

The purpose of reproducible research in public health is to ensure that results are **reliable** and *accurate*. If someone can reproduce your results, it means your methodology was correct and the data can be verified. Without reproducibility, the credibility/reliability of results become uncertain.

Advantages of Quarto:

-   Quarto updates live, so you don't have to constantly go between your code and the word document

-   Reduces the presence of errors because you don't have to constantly switch between the code editor and the text

## Code Chunk Customization

Create an R code chunk below with chunk option `#| echo: true` that creates a vector of systolic blood pressure readings: `120, 134, 118, 142, 128` and calculates their standard deviation using `sd()`.

```{r}
#| echo: true
# Write your code here:
systolic_blood_pressure <- c(120, 134, 118, 142, 128)
sd(systolic_blood_pressure)
```

------------------------------------------------------------------------

# Cycle 3: Git & GitHub Workflow

Confirm the following steps with your partner:

-   [x] You have initialized Git in your local R project.
-   [x] You have configured your Git identity in RStudio (`git config --global user.name` and `user.email`).
-   [x] You rendered this `.qmd` document to `.html` without any errors.

------------------------------------------------------------------------

# Cycle 4: AI Corner — Verification Exercise

An AI assistant was asked to write R code to compute the area of a circle with radius $r = 7.5\text{ cm}$.

The AI generated this code:

```{r ai-generated-code}
# AI-generated code:
radius <- 7.5
area <- 3.14 * (radius * 2)
```

### Questions:

1.  Is the AI-generated code correct? What error did it make?

    The AI-generated code is incorrect because the formula for Area is pi \* r\^2, and the AI-generated code did 3.14 \* (r \* 2), which would yield an incorrect area value because it does not use the correct value for pi, and also multiplies the radius by 2 instead of squaring the radius.

2.  Correct the code below using R's built-in constant `pi` and proper formula ($\text{Area} = \pi r^2$):

```{r corrected-code}
# Corrected code:
radius <- 7.5 
area <- pi * (radius^2)
area 
```

------------------------------------------------------------------------

## Submission Checklist

-   [x] Both your name and your partner's name are in the `author` field at the top.

-   [x] Click **Render** in RStudio and verify the HTML looks clean.

-   

    ## \[ \] Commit and push your changes to GitHub.
